Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-03 Master Thesis/synopsis.tex
Fix map label
Fix argument in 2.6 and 3.1

\begin{proposition}\label{2.6}
Let $K_0/K$ be an algebraic extension.
\begin{enumerate}[(i)]
\item Suppose there exists a prime number $\ell$ with $\ell^\infty\nmid[K_0:K]$ and such that for every algebraic extension $L/K_0$ we have: 
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K_0]\\
0, & \text{otherwise}
\end{cases} \]
Then there exists a unique henselian $w\in|\operatorname{Spv}(K)|$.
\item Assume that (i) is satisfied for two different prime numbers. Then $w$ is a defectless valuation.
\end{enumerate}
\end{proposition}

\begin{proof}
\begin{enumerate}[(i)]
\item Since $\mathrm{H}^3(K_0, \mathbb{Z}/\ell\mathbb{Z})\neq 0$ by assumption, according to Lemma \ref{2.4} we can find a valuation $w_0\in|\operatorname{Spv}(K_0)|$ such that:
\[ \mathrm{H}^3((K_0)_{w_0}, \mathbb{Z}/\ell\mathbb{Z}) \neq 0 \]
By requirement $\ell^\infty \nmid [(K_0)_{w_0} : K_0]$. By replacing $K_0$ by $(K_0)_{w_0}$, we can assume that $w_0$ is henselian.

@@

Now $(K_0)_{w_1}$ is henselian both with respect to $w_1$ and with respect to the extension of $w_0$. Lemma \ref{2.2} yields $\operatorname{cd}(G_{(K_0)_{w_1}})\leq 2$, a contradiction. The uniqueness of $w$ is again a consequence of Lemma \ref{2.2}.
\item Choose the prime $\ell$ different from $\operatorname{char} k(w)$ so that (i) holds. According to Lemma \ref{2.5} we have:
\[\operatorname{rrank}(w)=\dim_\mathbb{Q}(\Gamma_w\otimes_\mathbb{Z}\mathbb{Q} )\geq \dim_{\mathbb{F}_\ell}(\Gamma_w /\ell\Gamma_w) = 2 = \dim(K)\qedhere\]
\end{enumerate}
\end{proof}

@@

\begin{proof}
Let $\ell_1,\ell_2$ be two different prime numbers different from $\operatorname{char} k(w)$ and let $E_0 = E(\mu_{\ell_1\ell_2})$. Then, by Lemma \ref{2.5}, $E_0/E$, and also $E'_0/E'$, satisfy the condition of Proposition \ref{2.6} (ii). Thus there exists a unique henselian defectless valuation $w'_{E'}$ of $E'$, so $E'$ contains a henselization $K'_{w'}$ of $w' = w'_{E'}|_{K'}$. By applying the same argument to $\sigma^{-1}$ we find that $E' = K'_{w'}$.

@@

\begin{lemma}\phantomsection\label{4.3}
\begin{enumerate}[(i)]
\item $\varphi_n$ induces an isomorphism $\varphi_n: U_{P'}/n\to U_P/n$.
\item The following diagram commutes:
\[ \begin{tikzcd}
U_{P'}/n \rar["\varphi_n"] \dar["r_{P'}/n"'] & U_P/n \dar["r_P/n"]\\
k(P')^\times/n \rar["\varphi_n"'] & k(P)^\times/n
\end{tikzcd} \]
\end{enumerate}
\end{lemma}

>>>>

\begin{proposition}\label{2.6}
Let $K_0/K$ be an algebraic extension. Suppose there exists a prime number $\ell\neq 2$ with $\ell^\infty\nmid[K_0:K]$ and such that for every algebraic extension $L/K_0$ we have: 
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K_0]\\
0, & \text{otherwise}
\end{cases} \]
Then there exists a unique henselian $w\in|\operatorname{Spv}(K)|$.
\end{proposition}

\begin{proof}
Since $\mathrm{H}^3(K_0, \mathbb{Z}/\ell\mathbb{Z})\neq 0$ by assumption, according to Lemma \ref{2.4} we can find a valuation $w_0\in|\operatorname{Spv}(K_0)|$ such that:
\[ \mathrm{H}^3((K_0)_{w_0}, \mathbb{Z}/\ell\mathbb{Z}) \neq 0 \]
By requirement $\ell^\infty \nmid [(K_0)_{w_0} : K_0]$. By replacing $K_0$ by $(K_0)_{w_0}$, we can assume that $w_0$ is henselian.

@@

Now $(K_0)_{w_1}$ is henselian both with respect to $w_1$ and with respect to the extension of $w_0$. Lemma \ref{2.2} yields $\operatorname{cd}(G_{(K_0)_{w_1}})\leq 2$, a contradiction. The uniqueness of $w$ is again a consequence of Lemma \ref{2.2}.
\end{proof}

@@

\begin{proof}
Let $\ell_1,\ell_2$ be two different odd prime numbers different from $p=\operatorname{char} k(w)$ and let $E_0 = E(\mu_{\ell_1\ell_2})$. Since $K$ is finitely generated, by Lemma \ref{rank-dimension-inequality}, $k(w)$ and $\Gamma_w$ are finitely generated as well. In particular we have $\ell_i^\infty \nmid [k(w):\mathbb{F}_p]$ and, since $\Gamma_w$ is torsion free, $\dim_{\mathbb{F}_{\ell_i}}(\Gamma_w/{\ell_i}\Gamma_w) = \dim_{\mathbb{Q}}(\Gamma_w\otimes_{\mathbb{Z}}\mathbb{Q})= \operatorname{rrank}(w)=2$ for $i=1,2$.

By Lemma \ref{2.5}, $E_0/E$, and also $E'_0/E'$, satisfy the condition of Proposition \ref{2.6} for both $\ell_1$ and $\ell_2$. Thus there exists a unique henselian valuation $w'_{E'}\in|\operatorname{Spv}(E')|$. Choose a prime $\ell\neq \operatorname{char} k(w'_{E'})$ from $\ell_1,\ell_2$. Since $\mathrm{H}^3(E',\mathbb{Z}/\ell\mathbb{Z})\cong \mathrm{H}^3(E,\mathbb{Z}/\ell\mathbb{Z})\neq 0$, we get from Lemma \ref{2.5}:
\[\operatorname{rrank}(w'_{E'}) = \dim_{\mathbb{Q}}(\Gamma_{w'_{E'}}\otimes_\mathbb{Z}\mathbb{Q}) \geq \dim_{\mathbb{F}_\ell}(\Gamma_{w'_{E'}}/\ell\Gamma_{w'_{E'}})=2=\dim(E')\]
which makes $w'_{E'}\in\operatorname{Spv}^{(2)}_\text{d}(E')$. Now $E'$ contains a henselization $K'_{w'}$ of $w' = w'_{E'}|_{K'}\in\operatorname{Spv}^{(2)}_\text{d}(K')$. By applying the same argument to $\sigma^{-1}$ we find that $E' = K'_{w'}$.

@@

\begin{lemma}\phantomsection\label{4.3}
\begin{enumerate}[(i)]
\item $\varphi_n$ induces an isomorphism $\varphi_n: U_{P'}/n\to U_P/n$.
\item The following diagram commutes:
\[ \begin{tikzcd}
U_{P'}/n \rar["\varphi_n"] \dar["r_{P'}/n"'] & U_P/n \dar["r_P/n"]\\
k(P')^\times/n \rar["\varphi_P/n"'] & k(P)^\times/n
\end{tikzcd} \]
\end{enumerate}
\end{lemma}
